---
title: "WeatherModelSepPop"
author: "GGMFS Team"
date: "21/04/2021"
output: html_document
---

This script calculates growing degree days (GDD) using weather data from 6_Import_NOAAweather.R then interpolates GDD for each population

## Data prep and function load
```{r}
## Load functions
# Used for spatial interpolation of GDD for each population
library(fields) 
# Used to impute missing data points for some weather stations in some days
library(zoo) 
library(dplyr)
library(tidyr)

## Load data
PopData<-read.csv("../Population_Means_wClimate.csv")
StnData<-read.csv("../WeatherRawData/NOAAStationData.csv")
```

## Calculate GDD for 20 closest stations for each Pop
For each population in the dataset, find nearby stations in NOAAData, load station growing degrees for each day (GD).
NOTE: GD = (TMAX+TMIN)/20-5 and set GD=0 if GD<0
NOTE: /20 instead of 2 because measured in 1/10 degrees

PRCP - precipitation (1/10mm)
SNWD - snow depth (mm)
TMAX - max temp (1/10 deg C)
TMIN - min temp (1/10 deg C)
WESD - water equivalent of snow on ground (1/10mm) 
```{r}
# Standard Growing-Degree Days, as above
PopData$GDD<-NA 
# Number of days above 5oC
PopData$GDays<-NA 
Cntr<-0

if(!file.exists("../Population_Means_wClimate_wGDD.csv"){
  
  # For each year: 
  for(year in (2009:2013)){
    
    # Open file with GD data
    GDFilePath<-paste0("../WeatherRawData/NOAAStnsClose",year,".csv") 
    GDData<-read.csv(GDFilePath)
  
    # Cycle through pop_codes sampled in same year as GD data
    for(Pop in PopData$Pop_Code[PopData$Year==year]){ 
      Cntr<-Cntr+1
      Tmr<-Sys.time()
      # Record day of sampling
      SDay<-PopData$Day[PopData$Pop_Code==Pop]
      # Find names of nearby stations (stations near sampled populations)
      LocStns<-paste(unique(StnData$StationID[StnData$Pop_Code==Pop & StnData$Measure=="TMAX"]))
      
      # Subset GDData to include only nearby stations, and data from Jan 1 to the day of sampling
      PopGDData<-GDData[GDData$StationID %in% LocStns & GDData$Day<=SDay,]
      
      ###
      ## Remove stations with >10% missing data
      ###

      #Summarize sample size where NA are absent
     NaAbsent<- PopGDData %>% group_by(StationID) %>% drop_na(GD) %>% summarize(NaAbsent=n())
      #Summarize sample size where NA are present
     NaPresent<- PopGDData %>% group_by(StationID) %>% summarize(NaPresent=n())
     
     #Joining and determining % missing missing
     NaData<-NaPresent %>% left_join(NaAbsent,by="StationID") %>% mutate(PercentMissing=(NaPresent-NaAbsent)/NaPresent*100)
     
     #Isolating remaining populations.
     LocStns<-NaData %>% filter(PercentMissing<10) %>% select(StationID) %>% pull()
      
     ###
     #If there are less than 20 stations representing that population, then Station with >30% missing data will be removed to improve the sample size of the number of stations.
     ###
      if(length(LocStns)<20){
           LocStns<-NaData %>% filter(PercentMissing<30) %>% select(StationID) %>% pull()
      }
      
     #If there are more than 20 stations corresponding to the population...
      if(length(LocStns)>20){
        # Sort remaining stations by distance
        LocStns<-unique(StnData[StnData$Pop_Code==Pop & StnData$StationID %in% LocStns,c("StationID","Dist")])
        # Keep closest 20 stations 
        LocStns<-paste(LocStns$StationID[order(LocStns$Dist)][1:20])
      }
      
      # Re-Subset GDData now that we have screened for missing data and distance
      PopGDData<-GDData[GDData$StationID %in% LocStns & GDData$Day<=SDay,]
      
      # Make data frame to collect GDD, lat and long for each station
      GeoDat<-unique(StnData[StnData$StationID %in% LocStns,1:3])
      GeoDat$GDD<-NA
      GeoDat$GDays<-NA
      
      # Input missing data for each station
      for(Stn in LocStns){
        # Impute and then calculate GDD
        GeoDat$GDD[GeoDat$StationID==Stn]<-sum(na.approx(zoo(PopGDData$GD[PopGDData$StationID==Stn],0:SDay)))
        GeoDat$GDays[GeoDat$StationID==Stn]<-sum(na.approx(zoo(PopGDData$GD[PopGDData$StationID==Stn],0:SDay))>0)
        # Note: works even if no missing data
      }
      
      # Spatial interpolation of GDD
      if(nrow(GeoDat)>=10){ # Only interpolate if at least 10 stations available
        # Interpolate GDD for Pop
        tps<-Tps(x=GeoDat[,c("Longitude","Latitude")],Y=GeoDat$GDD,scale.type="unscaled") # see details in ?Tps  
        tpsD<-Tps(x=GeoDat[,c("Longitude","Latitude")],Y=GeoDat$GDays,scale.type="unscaled")
        PopData$GDD[PopData$Pop_Code==Pop]<-predict(tps,x=PopData[PopData$Pop_Code==Pop,c("Longitude","Latitude")],scale.type="unscaled")      
        PopData$GDays[PopData$Pop_Code==Pop]<-predict(tpsD,x=PopData[PopData$Pop_Code==Pop,c("Longitude","Latitude")],scale.type="unscaled")   
      }
      cat("***************\nIteration ",Cntr," of",length(unique(PopData$Pop_Code)),"\nYear: ",year,"\nPop: ",Pop,"\n",Sys.time()-Tmr,"seconds","\nGDD: ",PopData$GDD[PopData$Pop_Code==Pop],"\nGDays: ",PopData$GDays[PopData$Pop_Code==Pop],"\n***************")
      # clean up for next iteration of pop
      SDay<-LocStns<-PopGDData<-GDDs<-GeoDat<-tps<-NA 
    }
    # Clean-up for next iteration of year
    GDData<-GDFilePath<-NA 
  }
  
  # SAVE output
  write.csv(PopData,"../Population_Means_wClimate_wGDD.csv",row.names=F)

}
```



