---
title: "ClimaticNicheModels_PCReg"
author: "GGMFS Team"
date: "21/04/2021"
output: html_document
---


This script models climate effects using principal components regression

## Data prep and function load
```{r}
## Load packages
library(ggplot2)
library(gridExtra)
library(nlme)
library(vegan)

# Function to choose PCs based on statistical association with response variable
source("../functions/PCpick.R")
# Function for multiple regression with spatially autocorrelated error
source("../functions/mreg.R")
# Functions to generate barplots (barplot, theta function and angles function).
source("../functions/barplot_theta_angle.R")

## Load data
PopData<-read.csv("../Population_Means_wClimate_wGDD.csv")
PopData$Region<-factor(PopData$Region,levels=c("NorthAm","Europe"))
```

# Niche Model (with spatially autocorrelated error)

**Theory**: 
A redundancy analysis (RDA) with a single response vector (rather than matrix) simplifies to a multiple regression. But to prevent overfitting: principal components regression:
- 1. Standard PCA
- 2. Remove PCs with minimal eigenvalues
- 3. Regression on PC axes
- 4. Back-transformation of loadings from regression and PCA to identify key bioclim variables.
See Legendre & Legendre (2012) pg. 562-563.

## 1.  PCA of bioclim variables
```{r}
# Extract pcs from bio1-bio35 except 3 and 7 (which are calculated from other bios; bio 36:40 are PCs of the global data)
PCs<-PopData[,grep("bio",names(PopData))[1:35][c(-3,-7)]]
### NOTE: PCs 6, 22 and 30 are reversed to improve interpretation (e.g. bio5 is max temp; bio6 is min temp)
PCs$bio6<--PCs$bio6
PCs$bio22<--PCs$bio22
PCs$bio30<--PCs$bio30

# Replace original bioclim variables
PopData<-cbind(PopData[,grep("bio.*",names(PopData),invert=T)],PCs)

## Standardize bio variables to mean=0, sd=1
## Option not used (PCRegression uses correlation coefficients)
#for (col in names(PCs)[grep("bio",names(PCs))]){
#  PCs[,col]<-(PCs[,col]-mean(PCs[,col],na.rm=T))/sd(PCs[,col],na.rm=T)
#}

# Run principal components
bioPCs<-princomp(PCs,cor=T)
summary(bioPCs)
PCA<-1 # Choose PCs for plotting
PCB<-2
biplot(bioPCs,choices=c(PCA,PCB),xlabs=rep("",nrow(bioPCs$scores))) # Plot loadings of PCs 1&2
# Scale plot for adding points (note: approximation only!)
rrr<-apply(bioPCs$scores[,c(PCA,PCB)],2, range) 
rrr<-(abs(rrr)+1)*sign(rrr)
par(usr=as.vector(rrr)) 
# Add points, different colours for NA vs EU
points(bioPCs$scores[PopData$Region=="NorthAm",c(PCA,PCB)],col="#F5375166",pch=16)
points(bioPCs$scores[PopData$Region=="Europe",c(PCA,PCB)],col="#4FB0C666",pch=16)
```


## 2.  Reduce number of PC variables
```{r}
## Show PC loadings
plot(bioPCs)
summary(bioPCs) # Looks like 99% of variation in PC1-11

## Determine which PCs explain significant variation:
# For Methods 1 & 2 see Legendre & Legendre (2012) pg. 449
# Method 1: Broken stick model (Jackson 1993)
E0<-bstick(bioPCs) # Function from Vegan to calculate null eigenvalues
E1<-bioPCs$sdev^2

# Method 1a (Not Used): Retain PCs with eigenvalues > broken stick model (i.e. null model)
screeplot(bioPCs,bstick=T,type="lines")
keepPCs<-data.frame(bioPCs$scores[,E1>E0])
ncol(keepPCs) # Number of PCs retained

# Method 1b (Not Used): Retain PCs with cumulative eigenvalues > broken stick model
# Null variance
VarT0<-{}
# Observed variance
VarT<-{}
for(i in 1:length(E1)){
  VarT0<-c(VarT0,sum(E0[1:i]))
  VarT<-c(VarT,sum(E1[1:i]))
}
keepPCs<-data.frame(bioPCs$scores[,VarT>VarT0])
ncol(keepPCs) # Number of PCs retained

# Method 2 (Kaiser-Guttman criterion) (Not Used): Retain PCs with eigenvalues > average eigenvalue
keepPCs<-data.frame(bioPCs$scores[,E1>mean(E1)])
ncol(keepPCs) # Number of PCs retained

## Method 3: Retain PCs that are correlated with fitness components

# Keep Predictive PCs (The red line in plots denotes the simulated p value cut off where those below the line are not significant based on a permutation test with 1000 iterations.)
rFruits<-PCpick(Trait="MeanFruits",logT=T)
summary(lm(Resp~.,data=rFruits))

rTotalDens<-PCpick(Trait="TotalDens",logT=T)
summary(lm(Resp~.,data=rTotalDens))

rRosRatio<-PCpick(Trait="RosRatio",logT=F)
summary(lm(Resp~.,data=rRosRatio))

rPopSize<-PCpick(Trait="Pop_Size",logT=T)
summary(lm(Resp~.,data=rPopSize))

# Number of PCs retained:
length(grep("PC",names(rFruits)))
length(grep("PC",names(rTotalDens)))
length(grep("PC",names(rRosRatio)))
length(grep("PC",names(rPopSize)))

### NOTE: there is also a null expectation! (but bootstrap more reliable for non-normal data)
## See probability density function for Pearson product-moment correlation:
## http://en.wikipedia.org/wiki/Pearson_product-moment_correlation_coefficient
```


## 3. Regression on PC axes (with spatially autocorrelated errors)
```{r}
# Fruits
Fruits<-mreg(Data=rFruits)
qplot(fit,obs,data=Fruits$fit,colour=Region)+geom_smooth(method="lm")+theme_classic()
summary(lm(obs~fit,data=Fruits$fit))

# Total Density
TotalDens<-mreg(Data=rTotalDens)
qplot(fit,obs,data=TotalDens$fit,colour=Region)+geom_smooth(method="lm")+theme_classic()
summary(lm(obs~fit,data=TotalDens$fit))

# Rosette Ratio
RosRatio<-mreg(Data=rRosRatio)
qplot(fit,obs,data=RosRatio$fit,colour=Region)+geom_smooth(method="lm")+theme_classic()
summary(lm(obs~fit,data=RosRatio$fit))

# Population Size
PopSize<-mreg(Data=rPopSize)
qplot(fit,obs,data=PopSize$fit,colour=Region)+geom_smooth(method="lm")+theme_classic()
summary(lm(obs~fit,data=PopSize$fit))
```



# 4. Back-transformation of loadings from regression and PCA to identify key bioclim variables
```{r}
## Details in Legendre & Legendre (2012) pp 562-563
U<-as.data.frame(loadings(bioPCs)[,1:ncol(loadings(bioPCs))]) # Eigenvectors from original PCA
names(U)<-paste0("PC",1:ncol(loadings(bioPCs)))
## Calculate coefficients of original bioclim variables
# 1. Calculate matrix containing coefficients from PC regression (i.e. each ROW is a PC)
C<-data.frame(Fruits=1:ncol(loadings(bioPCs))*0,TotalDens=1:ncol(loadings(bioPCs))*0,
              RosRatio=1:ncol(loadings(bioPCs))*0,PopSize=1:ncol(loadings(bioPCs))*0)
C$Fruits[as.numeric(gsub("PC","",row.names(Fruits$coef)))]<-Fruits$coef$coef
C$TotalDens[as.numeric(gsub("PC","",row.names(TotalDens$coef)))]<-TotalDens$coef$coef
C$RosRatio[as.numeric(gsub("PC","",row.names(RosRatio$coef)))]<-RosRatio$coef$coef
C$PopSize[as.numeric(gsub("PC","",row.names(PopSize$coef)))]<-PopSize$coef$coef

# 2. Multiply U by each column in C to obtain coefficients for original bioclim variables (i.e. each ROW is a BIOCLIM variable)
B<-data.frame(Fruits=as.matrix(U) %*% C$Fruits,
              TotalDens=as.matrix(U) %*% C$TotalDens,
              RosRatio=as.matrix(U) %*% C$RosRatio,
              PopSize=as.matrix(U) %*% C$PopSize)

write.csv(B,"ClimLoadings.csv",row.names=T)

# Use B to calculate predictions for each performance measurement, and save to main dataset
PopData$PCFruits<-apply(scale(PCs),1,function(x) sum(x*B$Fruits))
PopData$PCTotalDens<-apply(scale(PCs),1,function(x) sum(x*B$TotalDens))
PopData$PCRosRatio<-apply(scale(PCs),1,function(x) sum(x*B$RosRatio))
PopData$PCPopSize<-apply(scale(PCs),1,function(x) sum(x*B$PopSize))

qplot(y=log(MeanFruits+1),x=PCFruits,colour=Region,data=PopData)+theme_classic()+geom_smooth(method="lm",se=F)
qplot(y=log(TotalDens+1),x=PCTotalDens,colour=Region,data=PopData)+theme_classic()+geom_smooth(method="lm",se=F)
qplot(y=RosRatio,x=PCRosRatio,colour=Region,data=PopData)+theme_classic()+geom_smooth(method="lm",se=F)
qplot(y=log(Pop_Size+1),x=PCPopSize,colour=Region,data=PopData)+theme_classic()+geom_smooth(method="lm",se=F)
```


# Generating histograms
```{r}
# Calculate correlations for each 'PC'
rPCs<-data.frame(Fruits=apply(PCs,2,function(x) cor(x,PopData$PCFruits)),
                     TotalDens=apply(PCs,2,function(x) cor(x,PopData$PCTotalDens)),
                     RosRatio=apply(PCs,2,function(x) cor(x,PopData$PCRosRatio)),
                     PopSize=apply(PCs,2,function(x) cor(x,PopData$PCPopSize)))

# Compare similarity (i.e. angles, in degrees) between matrices
### TO DO: Make function to input B (coefficients) or rPCs (R)

# Show similarity:
# Angles based on coefficients estimated from PC regression
# Should be interpreted cautiously when variables are inter-correlated
angles(B)
# Angles based on correlation coefficients
angles(rPCs)

## Plot eigenvectors using colour-coded vectors
categ<-data.frame(cat=1:35*NA,clr=1:35*NA)
row.names(categ)<-paste0("bio",1:35)
# Temperature
categ$cat[c(1,5,6,8,9,10,11)]<-"T"
categ$clr[c(1,5,6,8,9,10,11)]<-"#D72638"
categ$cat[c(2,4)]<-"T_Var"
categ$clr[c(2,4)]<-"#D7263866"

# Precipitation
categ$cat[c(12,13,14,16,17,18,19)]<-"Pcp"
categ$clr[c(12,13,14,16,17,18,19)]<-"#51BBFE"
categ$cat[15]<-"Pcp_Var"
categ$clr[15]<-"#51BBFE66"

# Moisture
categ$cat[c(28,29,30,32,33,34,35)]<-"Moist"
categ$clr[c(28,29,30,32,33,34,35)]<-"#8FF7A7"
categ$cat[31]<-"Moist_Var"
categ$clr[31]<-"#8FF7A766"

# Radiation
categ$cat[c(20,21,22,24,25,26,27)]<-"Rad"
categ$clr[c(20,21,22,24,25,26,27)]<-"#F9C80E"
categ$cat[23]<-"Rad_Var"
categ$clr[23]<-"#F9C80E66"


# Plot coefficients (i.e. B)
barplot(Trait="Fruits",abs=F,type="Coef")
barplot(Trait="TotalDens",abs=F,type="Coef")
# Plot correlation coefficients
barplot(Trait="Fruits",abs=F,type="R")
barplot(Trait="TotalDens",abs=F,type="R")

# Show legend
barplot(Trait="Loading",type="Legend",abs=T)

# Save Population means with PCRegression Predictions
write.csv(PopData,"../PCData.csv",row.names=F)
```


